RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-31 casestudies.tex
yes

\section{Case Studies}

We present two example applications in RCloud based on synthetic data
and applications.\footnote{Unfortunately, we could not show production
notebooks that may contain proprietary or confidential information.}
The main point of the examples is to show key steps in the visual
analytics development and deployment process that we aimed to support.

\subsection{Stock Price Analysis\label{sec:stockvis}}

\begin{figure}
\includegraphics[width=\linewidth]{fig/casestudy1/casestudy1.pdf}
\caption{\label{sec:stockvis}Iterations of a stock ticker
  visualization, based on an example by Hadley Wickham. A
  simple, static visualization of the closing price of a single stock
  is progressively developed into a configurable display suitable
  for dashboarding, then into an interactive visualization for 
  comparing the volatility and volume of two stocks, and finally
  into an API for data access by other RCloud notebooks. The notebooks
  in this example can all be loaded as web pages. When a notebook
  corresponding to a function call is displayed a web page,
  its associated documentation is displayed.}
\end{figure}

Our first example is a sequence of visualizations of the performance
of financial stocks over a multi-year period. The first visualization
in this example uses ggplot2 and is due to Hadley Wickham. It reads
data provided by Yahoo! Finance as a web service, and shows the price
for a single trading symbol.

The webpage that produces that visualization includes a link back to
the generating source code, from which a user can fork the notebook
and, in this case, add a configurable ticker based on the URL of
the notebook. Notice that the notebook changes very little.

From there, the notebook author or a collaborator may decide to
provide convenient access to the data, apart from its visualization.
This is done by creating a notebook that defines a function.
That notebook then becomes a \emph{subroutine} for other notebooks,
and is version-controlled like other notebooks.

\todo{Interactive version.}

%% This notebook can then be called in an \emph{interactive}
%% visualization, which uses the Javascript

%% Easy to convert it into configurable entry point for dashboarding.

%% Now consider an analyst that wants to understand the price dynamics of
%% different stock attributes, different points in time, etc. For that,
%% interactive visualization is very attractive. describe dcplot and
%% notebook which uses it.

%% Then describe abstraction of stock-data fetching, to talk about
%% calling notebooks from other notebooks.

%% then call.r?

\subsection{text analysis\ref{sec:textvis}}

Same.

>>>>

\section{Case Study}

We present an example Stock Price Analysis application
in RCloud.\footnote{Unfortunately,
we could not include production notebooks containing proprietary or personal
information.} The example, although simplified, shows key steps in the
development and deployment process we aimed to support.

\begin{figure}
\includegraphics[width=\linewidth]{fig/casestudy1/casestudy1.pdf}
\caption{\label{fig:stockvis}Iterations of a stock ticker
  visualization, based on an example by Hadley Wickham. A
  simple, static visualization of the closing price of a single stock
  is progressively developed into a configurable display suitable
  for dashboarding, then into an interactive visualization for 
  comparing the volatility and volume of two stocks, and finally
  into an API for data access by other RCloud notebooks. The notebooks
  in this example can all be loaded as web pages. When a notebook
  corresponding to a function call is displayed a web page,
  its associated documentation is displayed.}
\end{figure}

The example shows a sequence of visualizations of the performance
of financial stocks over a multi-year period. The first visualization
in this example uses ggplot2 and was written by Hadley Wickham.
It reads data provided by Yahoo! Finance as a web service, and shows
the price for a single trading symbol.

The webpage that produces that visualization includes a link back to
the underlying source code. From this link, a user can fork the notebook.
In the example,a configurable ticker is added, based on the URL of
the notebook. The change to the notebook is very minor.

Next, the notebook author or a collaborator may decide to extend
the application to also provide convenient access to the data,
apart from its visualization.
This is done by creating a notebook that defines a function.
That notebook then becomes a \emph{subroutine} for other notebooks.
It can be invoked from interactive notebooks, such as dashboards.
The data access notebook is is version-controlled like other notebooks.

Finally, consider the situation where an analyst wants to understand
the price dynamics of stocks with respects to various attributes and
time ranges. For this, interactive visualization may be very helpful.
In our example, the analyst creates an interactive tool with N views.
\todo{Describe dcplot and notebook which uses it.}

% \subsection{Text analysis\label{sec:textvis}}
% Same.

### 2014-03-31 system.tex
yes

\todor{Browse/Search}. ``More like this'', and a custom webpage solely for browsing.

@@

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This has a variety of
advantages which we discuss \todo{where?}. On the other hand, there
are a variety of situations that might ask for a specific version of a
notebook. Rather than have the notebook designer decide ahead of time
which versions should be preserved, we embrace \emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \todo{cite}
and VisTrails's version tree \todo{cite}, where every change in the state of the
system is tracked.

Implementing the entire version-control subsystem would be a large
undertaking. Fortunately, we can build RCloud on top of Github's
\emph{gists}~\todo{cite}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation being the
use of text-only files in a single directory). The GitHub web-service
API provides most of the semantics we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

>>>>

\todo{Browse/Search}. ``More like this'', and a custom webpage solely for browsing.

@@

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This has a variety of
advantages which we discuss \todo{where?}. On the other hand, there
are a variety of situations that might ask for a specific version of a
notebook. Rather than have the notebook designer decide ahead of time
which versions should be preserved, we embrace \emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \cite{Jankun-Kelly:2007:MFV}
and VisTrails's version tree \cite{Callahan:2006:VVM}, where every change in the state of the system is tracked.

Implementing the entire version-control subsystem would be a large
undertaking. Fortunately, we can build RCloud on top of Github's
\emph{gists}~\cite{Github:2014:GG}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation being the
use of text-only files in a single directory). The GitHub web-service
API provides most of the semantics we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.
